---
title: "Cellphone results plot"
project: "Dental-pulp"
author: "Martin Loza"
date: "2022/08/23"
---

On this workflow I aim to plot the results from the dental-pulp cellphone analysis
for the Fibroblast and Endothelial cells

## Setup 

```{r setup, message = FALSE}
library(dplyr)
library(ggplot2)
library(here)
library(patchwork)
library(stringr)
library(chorddiag)
library(igraph)
library(circlize)

# Global variables
seed = 777
```

# Load data

Load results from cellphone

```{r}
# data <- read.table(file = here("Data/pvalues.txt"), header = TRUE, sep = "\t")
data <- read.table(file = here("Data/significant_means.txt"), header = TRUE, sep = "\t")
```

## Setup

### Filter columns 

Select Fibroblasts and Endothelial interactions. Retain gene_a and gene_b information

```{r}
n_col <- colnames(data)
idx_gene <- which(n_col %in% c("gene_a", "gene_b"))
idx_cells <- which(str_detect(string = n_col, pattern = "Fibroblast") &
                     str_detect(string = n_col, pattern = "Endothelial"))
keep_idx <- c(idx_gene, idx_cells)
data <- data[,keep_idx]
rm(n_col, idx_cells, idx_gene, keep_idx)
```

In the interactions we have a transmitter and a receptor. For now let's analyse as:
Transmitter = Fibroblasts
Receptor = Endothelial 

```{r}
colnames(data)
```
Then, we should remove the other columns

```{r}
keep_idx <- which(str_detect(colnames(data), pattern = ".Fibroblast", negate = TRUE))
data <- data[,keep_idx]
```

## Test
remove fibroblast 7 and 6

```{r}
keep_idx <- which(!(str_detect(colnames(data), pattern = "Fibroblast_6", negate = FALSE) |
                    str_detect(colnames(data), pattern = "Fibroblast_7", negate = FALSE)))
data <- data[,keep_idx]
data <- data %>% select(!(Fibroblast_6.Odontoblast: Fibroblast_7.Odontoblast))
```

### Filter rows

Filter rows with empty gene info
```{r}
rmv_idx <- union(which(data$gene_a == ""), which(data$gene_b == ""))
data <- data[-rmv_idx,]
rm(rmv_idx)
```

Filter rows not significant in the five Fibroblast groups

```{r}
rmv_idx <- which(rowSums(x = data[,3:ncol(data)], na.rm = TRUE) == 0)
data <- data[-rmv_idx,]
data
```

## Chord plot: Top 25

Sum the interaction in the fibroblast groups

```{r}
interaction = rowSums(data[,3:ncol(data)], na.rm = TRUE)
df <- data %>% mutate("interaction" = interaction) %>% select(gene_a, gene_b, interaction)

#Filter top interaction
df <- df %>% filter(interaction != 0) %>% top_n(n = 25, wt = interaction) %>% arrange(desc(interaction))
```

```{r}
#get adjacency list
g <- graph.data.frame(df[,1:2])
adj <- as.matrix(get.adjacency(g))

#Fill the adj matrix 
#for each interaction
for(i in seq_len(nrow(df))){
  current_interaction <- df[i,] 
  adj[current_interaction$gene_a, current_interaction$gene_b] <- current_interaction$interaction
}
```

```{r, fig.width=4}
adj2 <- adj[rev(rownames(adj)),]
arr <- data.frame(c(rep("MDK", 2), rep("PTN", 2)), c("LRP1", "PTPRZ1", "PTPRZ1", "PTPRS"), rep(1, 4))
arr2 <- data.frame(c(rep("MDK", 2), rep("PTN", 2)), c("LRP1", "PTPRZ1", "PTPRZ1", "PTPRS"), 1:4)
arr2 <- data.frame(c(rep("MDK", 2), rep("PTN", 2)), c("LRP1", "PTPRZ1", "PTPRZ1", "PTPRS"), rep(2, 4))
arrows <- data.frame(c(rep("MDK", 2), rep("PTN", 2)), c("LRP1", "PTPRZ1", "PTPRZ1", "PTPRS"), rep("black", 4))
  
pdf(file = here("chord_plot_Endothelial_220831.pdf"), width = 7, height = 7, bg = "white")

par(cex = 0.75)
circos.par(start.degree = 40, clock.wise = TRUE)

set.seed(111)
chordDiagram(adj2,annotationTrack = c("grid"),
             transparency = 0.25,
             # directional = 1,
             # direction.type = "arrows",
             # link.arr.col = arrows,
             # link.arr.length = 0.6, 
             preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(adj2))))))
circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
}, bg.border = NA)
circos.clear()
dev.off()
```



